Data Specialization for a Staged Lambda Calculus

### 2015-06-28 paper/splitting.tex
working on splitting

\textbf{\bbtwo\ term splitting.}  For any $\coltwo e A$ which reduces
to a residual $q$ (via $\diatwo e q$) and then a final value $v$ (via
$\reduce q v$), splitting $e$ (via the new relation~$\splittwosym$)
yields a pair of monostaged terms $\pipeS p l r$. ($p$ and $l.r$
contain all first- and second-stage computations in $e$ respectively.)
Executing $p$ in a first pass produces $b$ (via $\reduce p b$). Then,
in the second pass, $\reduce {[b/l]r} {v'}$, where $v'$ is the same as
$v$.  Since execution of the first pass serves to generate input for
the second pass, we say $p$ is a {\em precomputation} that produces a
{\em boundary value} ($b$) for the {\em resumer} ($l.r$).

@@

\textbf{\bbonem\ term splitting.}  For any $\colmix e A$ which reduces
to $\rtab \xi v$ (via $\redonesym$) , splitting $e$ (via the new
relation~$\splitonesym$) again yields the pair of monostage terms
$\pipeM c l r$. Evaluating $c$ in the first pass produces the pair
$(i,b)$, where the first element $i$ matches the first-stage results
contained in $v$.  For example, in the simple case where $\rtab \xi v$
contains entirely first-stage computation, such as:
\begin{lstlisting}
2`[]`1` 3`gr{`1`literalone+literaltwo`3`}`
\end{lstlisting}
the first-stage result is the value 3.  The second element of the pair
is the boundary value used as input to the resumer in second pass
evaluation as $\reduce {[b/l]r} {v'}$, where $v'$ is the same as
result produced by second-stage computations in $e$.  For example, in
the case where $\rtab \xi v$ contains entirely second-stage
computation, such as:
\begin{lstlisting}
2`[yhat |-> 2+3]`1` next{`2`yhat`1`}`
\end{lstlisting}
the second-stage result is the value 5. Since in general residual
tables and partial values describe a mixture of both first and second
stage computations, in the subsequent subsection we define an
operation, called {\em masking}, that separates residual tables and
partial values into their first- and second-stage results. We then
describe the algorithm for splitting both \bbonem\ terms
($\splitonesym$) and \bbtwo\ terms ($\splittwosym$).

>>>>

\textbf{\bbtwo\ term splitting.}  For any $\coltwo e A$
which reduces to the residual $\exv q$ (via repeated $\stepsym \bbtwo$),
splitting $e$ (via the new relation~$\splittwosym$)
yields a pair of monostaged terms $\pipeS p l r$, where $p$ and $l.r$
contain all first- and second-stage computations in $e$ respectively.

First pass evaluation reduces $p$ to $b$,
which is plugged in for $l$ to produce $[b/l]r$.
Correctness of splitting states that this should be equivalent 
to the residual that would have been produced by direct evaluation.
That is, $[b/l]r \equiv q$.

Since execution of the first pass serves to generate input for
the second pass, we say $p$ is a {\em precomputation} that produces a
{\em boundary value} ($b$) for the {\em resumer} ($l.r$).

@@

\textbf{\bbonem\ term splitting.}  For any $\colmix e A$ which reduces
to $\exv v$ (via repeated $\stepsym \bbonem$), 
splitting $e$ (via the new relation~$\splitonesym$) again 
yields the pair of monostage terms $\pipeM c l r$. 
Evaluating $c$ in the first pass produces the pair
$(i,b)$, where the first element $i$ matches the first-stage results
contained in $v$.  For example, in the simple case where $e$
contains entirely first-stage computation, such as:
\begin{lstlisting}
1` 3`gr{`1`literalone+literaltwo`3`}`
\end{lstlisting}
the first-stage result is the value 3.  The second element of the pair, $b$,
is the boundary value used as input to the resumer in second pass
evaluation as $[b/l]r \equiv q$, where $q$ is the same as
result produced by second-stage computations in $e$.  For example, in
the case where $\rtab \xi v$ contains entirely second-stage
computation, such as:
\begin{lstlisting}
2`[yhat |-> 2+3]`1` next{`2`yhat`1`}`
\end{lstlisting}
the second-stage result is the value 5. Since in general residual
tables and partial values describe a mixture of both first and second
stage computations, in the subsequent subsection we define an
operation, called {\em masking}, that separates residual tables and
partial values into their first- and second-stage results. We then
describe the algorithm for splitting both \bbonem\ terms
($\splitonesym$) and \bbtwo\ terms ($\splittwosym$).
